---
title: "Florida Youth Suicide"
author: "Andriy V. Koval & Kyle Belanger"
date: "April 30, 2020"
output:
  html_document:
    theme: simplex
    toc: true
    toc_depth: 3
    keep_md: false
    toc_float: true
    code_folding: hide
---

<!-- These two chunks should be added in the beginning of every .Rmd that you want to source an .R script -->
<!--  The 1st mandatory chunck  -->
<!--  Set the working directory to the repository's base directory -->
```{r, echo=F, message=F} 
#Don't combine this call with any other chunk 
# cat("Working directory: ", getwd()) # check where you are
  library(knitr)
# Rmd is in "./reports/reportA/reportA.Rmd", but now the root is "./"
  knitr::opts_knit$set(root.dir='../../') # "go up" from the folder where this .Rmd is save until you arrive at the root/working directory
```

<!--  The 2nd mandatory chunck  -->
<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F}
# set options shared by all chunks
opts_chunk$set(
  results='show', 
  attr.source = ".numberLines",
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4, 
  fig.width = 5.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/',     
  dev = "png",
  dpi = 400
  # fig.path = 'figure_pdf/',     
  # dev = "pdf"#,
  # dev.args=list(pdf = list(colormodel = 'cmyk'))
)
echoChunks <- TRUE
options(width=120) #So the output is 50% wider than the default.
# connect to the file with the underlying R script  
read_chunk("./analysis/florida-suicide-10_24/fl-suicide-10_24.R")

```

# Abstract

<!-- NEED TO UPDATE  -->

This blogpost explores youth (ages 10 - 24) suicide trends in Florida between 2006 and 2017, examining the effects of race and gender. The post answers the following questions with respect to population in focus:  

> Q1 - What is the overall trajectory of youth suicides in FL between 2006 and 2017? 

> Q1b - How do trends in population growth, suicide count and suicide rate compare among age group? 

> Q2 - How do means of suicide compare in each year? 

> Q3 - How do trends compare across the means of suicide?

> Q4 - How do trends compare by race, sex, and means of means of suicide?

# Environment

___This section contains technical information for deeper analysis and reproduction. Casual readers are invited to skip it.___   

Packages used in this report. 
<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-packages, echo=echoChunks, message=FALSE, warning = FALSE}
```
Script dependencies.
```{r load-sources, echo=echoChunks, message=FALSE, warning = FALSE}
```
Definitions of global object (file paths,  factor levels, object groups ) used throughout the report. 
```{r declare-globals, echo=echoChunks, results='show', message=FALSE}
```
Definitions of custom functions.  
```{r declare-functions, echo=echoChunks, results='show', message=FALSE}
```

# Data  

The data comes from Florida Health Charts and contains suicide mortality between 2006 and 2017, broken down by suicide means, county, sex, age, and race. The groomed dataset is [available for download here](https://github.com/dss-hmi/suicide-prevention-2019/blob/b225979bad5e4728d7a594fd455ab929d3ccafb0/data-public/derived/9-population-suicide.csv).  

```{r load-data, echo=echoChunks, results='show', message=FALSE}
```


## Data Tweaks
This dataset already comes  to this report well-groomed, so only some minor housekeeping tweaks are necessary to make graphing more convenient. The data is also filtered to only include individuals between ages 10 and 24.
```{r tweak-data-1, echo=echoChunks, results='show', message=FALSE,cache=TRUE}
ds0 %>% glimpse()
```

The unit of analysis of this dataset is defined by the first 6 variables: `county`, `year`, `sex`, `race`, `ethnicity`, `age_group`. For each unit, there are two measures: the number of people (`n_population`) and the number of observed suicide events (`n_suicides`). The latter is broken down by 7 means of suicide. For convenience, we create a combined variable `race_ethnicity`.

```{r}
ds0 %>% group_by(race, ethnicity, race_ethnicity) %>% count() %>% select(-n)
```

# Q1 - Overall Trend

To begin, let us answer the most fundamental question: 

> Q1 - What is the overall trajectory of youth suicides in FL between 2006 and 2017? 

```{r q1-1, echo=TRUE, fig.height=6, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

> Q1b - How do trends in population growth, suicide count and suicide rate compare among age group? 

```{r q1b-1, echo=TRUE, fig.height=7, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

The `10-14` age group witnesses quite a dramatic decrease in the `One out of` measure, dwarfing the scale for the other two age groups. We recreate it separately:

```{r q1b-2, echo=TRUE, fig.height=3, fig.width=8, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

To explore whether certain age transitions are associated with increased suicidality, we plot the distribution of suicide events within each year of age among the observed years (2006 - 2018). 
```{r q1b-3, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE, warning=FALSE}
```
We see a marked spike in deaths during the 17 to 18 transition. 
It appears the majority of deaths in 10-14 groups could be attributed to 13 and 14 year olds. 
There appears to be a plateau following the age 25 for the next 15 years, which makes theoretical sence: after the turmoils of coming out of age and young adulthood, there life acquires a more stable character. 

Also note, that when examining the trends within the `10-24` age group, the relationship bewtween the age and the suicide mortality is very well summarized by a linear model ($R^2 = .85$)
```{r q1b-4, echo=TRUE, fig.height=7, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

# Q2 - Yearly Breakdowns

> Q2 - How do means of suicide compare in each year? 

```{r q2-1, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


```{r q2-2, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```




# Q3 - Means across Years 

> Q3 - How do trends compare across the means of suicide?

```{r q3-1, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

```{r q3-2, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


When ignoring race and ethnicity for age group 10-24 average increase of suicide mortality from gun (+0.1) per year is simaliar to average increase from non-gun means (+0.105)

# Q4 - Sex and Ethnicity

> Q4 - How do trends compare by race, sex, and means of means of suicide?

```{r q4-1, echo=TRUE, fig.height=7, fig.width=11, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


```{r q4-2, echo=TRUE, fig.height=7, fig.width=11, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```




session information  
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
} 
```

