---
title: 'GLS Activity: Maps'
output:
  html_document:
    always_allow_html: yes
    df_print: kable
    highlight: tango
    keep_md: yes
    table: kable
    theme: spacelab
    toc: yes
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  word_document:
    highlight: haddock
    toc: yes
  pdf_document:
    toc: yes
---

<!-- These two chunks should be added in the beginning of every .Rmd that you want to source an .R script -->
<!--  The 1st mandatory chunck  -->
<!--  Set the working directory to the repository's base directory -->
```{r, echo=F, message=F} 
#Don't combine this call with any other chunk 
# cat("Working directory: ", getwd()) # check where you are
  library(knitr)
# Rmd is in "./reports/reportA/reportA.Rmd", but now the root is "./"
  knitr::opts_knit$set(root.dir='../../') 
```

<!--  The 2nd mandatory chunck  -->
<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F}
# set options shared by all chunks
report_render_start_time <- Sys.time()
opts_chunk$set(
  results='show', 
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 7, 
  fig.width = 12, 
  out.width = "900px",
  fig.path = 'figure_rmd/',     
  dev = "png",
  dpi = 400,
  # fig.path = 'figure_pdf/',     
  # dev = "pdf"#,
  # dev.args=list(pdf = list(colormodel = 'cmyk'))
  class.source = "numberLines lineAnchors", 
  class.output = c("numberLines lineAnchors chunkout")
)
echoChunks <- FALSE
warningChunks <- FALSE
options(width=120) #So the output is 50% wider than the default.
# connect to the file with the underlying R script  
read_chunk("./analysis/gls-maps/gls-maps.R") 
```


# Set the scene 

Describe the working environment. 
<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-packages, echo=echoChunks, message=FALSE,warning=warningChunks}
```


<!-- Load the sources.  Suppress the output when loading sources. --> 
```{r load-sources, echo=echoChunks, message=FALSE,warning=warningChunks}
```


<!-- Load any Global functions and variables declared in the R file.  Suppress the output. --> 
```{r declare-globals, echo=echoChunks, results='show', message=FALSE,warning=warningChunks}
```

<!-- Declare any global functions specific to a Rmd output.  Suppress the output. --> 
```{r, echo=echoChunks, message=FALSE,warning=warningChunks}
#Put code in here.  It doesn't call a chunk in the codebehind file.
```

<!-- Load the datasets.   -->
```{r load-data, echo=echoChunks, results='show', message=FALSE,warning=warningChunks}
```

<!-- Inspect the datasets.   -->
```{r inspect-data, echo=echoChunks, results='show', message=FALSE,warning=warningChunks}
```

<!-- Tweak the datasets.   -->
```{r tweak-data, echo=echoChunks, results='show', message=FALSE,warning=warningChunks}
```


# I Summary view

```{r define-map-1 , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 7, out.width = "960px"} 
```
## Prof
```{r map-1-professionals , echo=echoChunks, results='show', message=T,warning=T, fig.width = 17, fig.height = 17, out.width = "960px"} 
```

## Comm
```{r map-1-community , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 17, out.width = "960px"} 
```

```{r peer-counties-map , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 17, out.width = "960px"} 
```

```{r dev-1 , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 7, out.width = "960px"} 
```

```{r dev-2 , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 7, out.width = "960px"} 
```


```{r recap-1 , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 7, out.width = "960px"} 
```

```{r recap-2 , echo=echoChunks, results='show', message=FALSE,warning=warningChunks, fig.width = 17, fig.height = 7, out.width = "960px"} 
```




Session Information {#session-info}
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
}
```
</details>

```{r session-duration, echo=FALSE}
report_render_duration_in_seconds <- round(as.numeric(difftime(Sys.time(), report_render_start_time, units="secs")))
```

Report rendered by `r Sys.info()["user"]` at `r strftime(Sys.time(), "%Y-%m-%d, %H:%M %z")` in `r report_render_duration_in_seconds` seconds.


