---
title: "Florida Demographic Growth"
author: "Andriy Koval"
date: "March 27, 2020"
output:
  html_document:
    theme: simplex
    toc: true
    toc_depth: 3
    keep_md: true
    toc_float: true
    code_folding: hide
---

<!-- These two chunks should be added in the beginning of every .Rmd that you want to source an .R script -->
<!--  The 1st mandatory chunck  -->
<!--  Set the working directory to the repository's base directory -->
```{r, echo=F, message=F} 
#Don't combine this call with any other chunk 
# cat("Working directory: ", getwd()) # check where you are
  library(knitr)
# Rmd is in "./reports/reportA/reportA.Rmd", but now the root is "./"
  knitr::opts_knit$set(root.dir='../../../') # "go up" from the folder where this .Rmd is save until you arrive at the root/working directory
```

<!--  The 2nd mandatory chunck  -->
<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F}
# set options shared by all chunks
opts_chunk$set(
  results='show', 
  attr.source = ".numberLines",
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4, 
  fig.width = 5.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/',     
  dev = "png",
  dpi = 400
  # fig.path = 'figure_pdf/',     
  # dev = "pdf"#,
  # dev.args=list(pdf = list(colormodel = 'cmyk'))
)
echoChunks <- TRUE
options(width=120) #So the output is 50% wider than the default.
# connect to the file with the underlying R script  
read_chunk("./analysis/blogposts/florida-suicide-count/fl-suicide-count.R") 
# read_chunk("../../../analysis/blogposts/florida-suicide-count/fl-suicide-count.R") 
```
# Abstract



# Environment
<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-packages, echo=echoChunks, message=FALSE}
```

 ```{r declare-globals, echo=echoChunks, results='show', message=FALSE}
 ```
 
```{r declare-functions, echo=echoChunks, results='show', message=FALSE}
 ```

# Data  

The data comes from Florida Health Charts and contains suicide mortality between 2006 and 2017, broken down by suicide means, county, sex, age, and race. The the dataset is [available for download here](). 
```{r load-data, echo=echoChunks, results='show', message=FALSE}
```


## Data Tweaks
This dataset already comes  well-gromed, so only some minor housekeeping tweaks are necessary to make graphing more convenient. 
```{r tweak-data-1, echo=echoChunks, results='show', message=FALSE}
ds0 %>% glimpse()
```

The unit of analysis of this dataset is defined by the first 6 variables: `county`, `year`, `sex`, `race`, `ethnicity`, `age_group`. For each unit, there are two measures: the number of people (`n_population`) and the number of observed suicide events (`n_suicides`). The latter is broken down by 7 means of suicide. For convenience, we create a combined variable `race_ethnicity`.

```{r}
ds0 %>% group_by(race, ethnicity, race_ethnicity) %>% count() %>% select(-n)
```

# Q1 - Overal trends

To begin, let us answer the most fundamental question

> What is the overall trajectory of suicides in FL between 2006 and 2017? 



```{r g1, echo=TRUE, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

session information
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
} 
```
