---
title: "Florida Suicide (1)"
author: "Andriy Koval"
date: "April 26, 2020"
output:
  html_document:
    theme: simplex
    toc: true
    toc_depth: 3
    keep_md: true
    toc_float: true
    code_folding: hide
---

<!-- These two chunks should be added in the beginning of every .Rmd that you want to source an .R script -->
<!--  The 1st mandatory chunck  -->
<!--  Set the working directory to the repository's base directory -->
```{r, echo=F, message=F} 
#Don't combine this call with any other chunk 
# cat("Working directory: ", getwd()) # check where you are
  library(knitr)
# Rmd is in "./reports/reportA/reportA.Rmd", but now the root is "./"
  knitr::opts_knit$set(root.dir='../../../') # "go up" from the folder where this .Rmd is save until you arrive at the root/working directory
```

<!--  The 2nd mandatory chunck  -->
<!-- Set the report-wide options, and point to the external code file. -->
```{r set_options, echo=F}
# set options shared by all chunks
opts_chunk$set(
  results='show', 
  attr.source = ".numberLines",
  message = TRUE,
  comment = NA, 
  tidy = FALSE,
  fig.height = 4, 
  fig.width = 5.5, 
  out.width = "550px",
  fig.path = 'figure_rmd/',     
  dev = "png",
  dpi = 400
  # fig.path = 'figure_pdf/',     
  # dev = "pdf"#,
  # dev.args=list(pdf = list(colormodel = 'cmyk'))
)
echoChunks <- TRUE
options(width=120) #So the output is 50% wider than the default.
# connect to the file with the underlying R script  
read_chunk("./analysis/blogposts/florida-suicide-1/fl-suicide-1.R")
# read_chunk("../../../analysis/blogposts/florida-suicide-count/fl-suicide-count.R")
```
# Abstract

This blogpost explores suicides  trends in Florida between 2006 and 2017, examining the difference by age, gender, and race. The post answers the following questions:  

> Q1 - What is the overall trajectory of suicides in FL between 2006 and 2017? 

> Q2 - How does change in suicide mortality differ by age? 

> Q3 - How do suicide trends differ by gender? 

> Q4 - How do suicide trends differ by gender and age group?

> Q5 - How do suicide trend differ by race?

# Environment
<!-- Load 'sourced' R files.  Suppress the output when loading packages. --> 
```{r load-packages, echo=echoChunks, message=FALSE}
```

 ```{r declare-globals, echo=echoChunks, results='show', message=FALSE}
 ```
 
```{r declare-functions, echo=echoChunks, results='show', message=FALSE}
```

# Data  

The data comes from Florida Health Charts and contains suicide mortality between 2006 and 2017, broken down by suicide means, county, sex, age group, and race. The the dataset is [available for download here](https://github.com/dss-hmi/suicide-prevention-2019/blob/b225979bad5e4728d7a594fd455ab929d3ccafb0/data-public/derived/9-population-suicide.csv).  

```{r load-data, echo=echoChunks, results='show', message=FALSE}
```


## Data Tweaks
This dataset already comes  well-gromed, so only some minor housekeeping tweaks are necessary to make graphing more convenient. 
```{r tweak-data-1, echo=echoChunks, results='show', message=FALSE,cache=TRUE}
ds0 %>% glimpse(90)
```

The unit of analysis of this dataset is defined by the first 6 variables: `county`, `year`, `sex`, `race`, `ethnicity`, `age_group`. For each unit, there are two measures: the number of people (`n_population`) and the number of observed suicide events (`n_suicides`). The latter is broken down by 7 means of suicide. In this report we focus on the __total number__ of suicide, regardless the means.  

For convenience, we create a combined variable `race_ethnicity`.

```{r}
ds0 %>% group_by(race, ethnicity, race_ethnicity) %>% count() %>% select(-n)
```

# Q1 - Overall Trend

To begin, let us answer the most fundamental question: 

> What is the overall trajectory of suicides in FL between 2006 and 2017? 

```{r g11, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

We exclude persons younger than `10` from our analysis. There were only a handful of cases of suicide among persons younger than 10. Here is the frequency of suicides for each age group: 

```{r g12, echo=TRUE, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```
As we see, there are only `4` suicide events for this age group. (The original data was retrieved in November 2019, as of May 2020, the FLDH has updated its mortality report and now includes only 2 cases of suicides among persons younger than 10: one in 2013 and another in 2016).  

We remove the youngest three age groups and redraw the graph with the same scale across the facets for a better comparison among age groups.

```{r g13, echo=TRUE, fig.height=7, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


# Q2 - Age

> How does change in suicide mortality differ by age? 

Note that age groups do not have the same definition: ages younger than `25` are grouped into 5-year bins, while older ages are grouped into 10-year bins. This makes it difficult to compare absolute counts across age groups. Also, we must account for the fact that demographic growth may be uneven across the groups, which it is:

```{r g21, echo=TRUE, fig.height=9, fig.width=12, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```
Notice that the age groups `10_14` and `15_19` have been ___declining___ by an average of `2,680` and `2,970` respectively, while `20_24` age group have been ___increasing___ by an average of `12,500` persons per year.  Due to uneven size, compared to other age groups, it's hard to make out trends for these three; we zoom in to take a closer look:  


```{r g21a, echo=TRUE, fig.height=3, fig.width=9, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

To compare the change in suicide mortality among different age groups, we compute an age-adjusted rate measure:

```{r g22, echo=TRUE, fig.height=9, fig.width=12, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

# Q3 - Sex 

> How do suicide trends differ by gender? 

Existing research suggests that suicide prevalence will differ substantially between the genders and we do observe it in Florida data.

```{r g31, echo=TRUE, fig.height=5, fig.width=10, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

# Q4 - Sex and Age

> How do suicide trends differ by gender and age group?


```{r g41, echo=TRUE, fig.height=9, fig.width=13, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

To help us compare the magnitute of suicide rate across age group, we redraw the graph with the same scale across the facets:

```{r g42, echo=TRUE, fig.height=9, fig.width=13, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

# Q5 - Race + Ethnicity

> How do suicide trend differ by race?

The data stored race and ethnicity into two separate variables, which we have combined to have well-populated groups
```{r}
ds0 %>% 
  dplyr::filter(year == 2017) %>% 
  group_by(race, ethnicity, race_ethnicity) %>%
  summarize(population_in_2017 = sum(n_population)) 
```
The differences in group size, combined with the rarity of suicide events make it difficult to compare the trends across groups on the same scale. To make comparison of the shapes of the trends easier, we plot each group and metric on individual scale:

```{r g51, echo=TRUE, fig.height=6, fig.width=12, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```

However, when working with `rate per 100,000` metric, which adjusts for group size, it makes sense to have both views: constrained to the same scale (top) and plotted on free scales (bottom) 

```{r g51a, echo=TRUE, fig.height=6, fig.width=12, out.width = "900px", message=FALSE, results='show',cache=TRUE}
```


Unfortunately, when we start breaking down the groups by more categories (`age_group`, `sex`), this problem gets worse. Therefore the plots exploring the relationships between age, sex, and race are created for each of the race category individually to to avoid comparing across drastically different scales.  The graphs below are not ideal: while allowing to explore pattern within each race group they fail to provide a direct comparison among them. A more nuanced visualization is needed for this combination of variables.   
```{r g52, echo=TRUE, fig.height=9, fig.width=23, out.width = "1200px", message=FALSE, results='asis',cache=TRUE, dpi = 200}
```

session information  
===========================================================================

For the sake of documentation and reproducibility, the current report was rendered in the following environment.  Click the line below to expand.

<details>
  <summary>Environment <span class="glyphicon glyphicon-plus-sign"></span></summary>
```{r session-info, echo=FALSE}
if( requireNamespace("devtools", quietly = TRUE) ) {
  devtools::session_info()
} else {
  sessionInfo()
} 
```
